The corona of a fully convective star with a near-polar flare

### 2023-08-28 src/tex/ms.tex
[showyourwork] overleaf sync

% Title
\title{Fully convective star that produced a near-polar flare shows a typical corona}

@@

% Abstract with filler text
\begin{abstract}
    In 2020, TESS observed a rapidly rotating M7 dwarf produce a flare at 81 deg latitude, the highest latitude flare located to date. This is in stark contrast to solar flares that occur much closer to the equator, typically below 30 deg. The circumstances that allow flares at high latitudes to occur are poorly understood. We studied five Sectors of TESS observations, and obtained 36 ks of XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to X-ray luminosity, coronal temperatures, and flaring activity. Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band. Combining the 18 flares found in TESS with the flare from XMM-Newton, we inferred that TIC 277 produces flares at or above this energy approximately once every 10 h. We conclude that TIC 277's X-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
Most M dwarfs hosts terrestrial planets with a sizable fraction orbiting at instellations where liquid water can exist~\citep{dressing2015occurrence, hardegree-ullman2019kepler, ment2023occurrence}. However, to be habitable in an Earth-like manner, merely the possibility of surface water in its liquid form does not suffice. Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere affects hospitability for life, for M dwarfs in particular~\citep{airapetian2020impact}. If the energetic (particle) flux is too high, the atmosphere may blow off~(XXXX). If not energetic enough, life may not emerge in the first place~\citep{rimmer2018origin}.

In M dwarfs, stellar activity evolution on the main sequence unfolds much slower than for a FGK star. While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{XXXX}. Moreover, the zone where liquid water can reside is much closer to these faint stars. The result is a planet that may be exposed to violent space weather conditions for a good fraction of the age of the universe.

Stellar space weather originates from the stellar corona. The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission. The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~(XXXX). Yet, brown dwarfs are regularly detected with energetic flares regardless of their low X-ray luminosity~(XXXX). How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{weber2023understanding}.%. An intermittent corona is conceivable, similarly XXXX. Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within, and consequently influences the complex process of flux emergence

\citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with TESS. They found four flares on four rapidly rotating ($P<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star. The flares were observed to rotate in and out view. The shape of this modulation together with known inclination (combining $P$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares. Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude. Their results indicate a preference for flares at high latitudes in these stars, as a chance finding was unlikely ($\sim 0.1\%$). If flares and associated particle eruptions indeed have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought. %Are high flaring latitudes  an effect of the dynamo at high rotation speeds? Or is there something special about these stars' coronae that allowed them to produce flares at these latitudes?

@@

TIC 277539431, TIC 277 for short, is one out of four stars with a high-latitude flare in its TESS observations~\citep[][Table \ref{tab:starparams}]{ilin2021giant}. In fact, it showed the highest latitude flare known to date at $\sim81\,$deg. TIC 277 is an M7 dwarf with an extremely short rotation period of 4.56 hours, detected in the rotational modulation of its light curve. While the fast rotation and late spectral type could be indicative of a brown dwarf~(), the detection of flares suggests the presence of a stellar corona~().%The star's age is unknown. \citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would, if not as high, be compatible with the NUV/FUV emission from younger open cluster members. 
%Is its high flaring latitude an effect of the dynamo at high rotation speeds? Or is there something special about this star's corona that led it to produce are flare at these latitudes? 
In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from the Transiting Exoplanet Survey Satellite~(TESS,~\citealt{ricker2015transiting}, Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}). We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star, brown dwarf, or something in between~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.

@@

\begin{table}
% \footnotesize
\centering
    \caption{TIC 277539431 (WISEA J105515.71-735611.3)}
    \begin{tabular}{ll}\hline 
         Parameter & Value  \\\hline
         Distance $d$ & $13.70\pm0.11\,$pc [2] \\
         % $G$ mag & $14.7189 \pm 0.0011$ [4]\\
         % $G_{BP}$ mag & \\
         % $G_{BP}$ mag & \\
         % $K_s$ mag & $9.666 \pm 0.024$ [3]\\
         % $J$ mag & $10.630 \pm 0.023$ [3] \\
         Spectral type & M7 [1]\\
         Effective temperature $T_{\rm eff}$ & $2680^{60}_{-50}\,$ K [3]\\
         % age & \\
         Rotation period $P_{\rm rot}$ & $273.618 \pm 0.007\,$min [1]\\
         Projected rotation velocity $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
         Inclination $i$ & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
         Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\
         Bol. luminosity $L_{\rm bol}$ & \Lbol \\\hline

@@

\subsection{XMM-Newton}
\label{sec:obs:xmm}
XMM-Newton is an X-ray telescope, which launched into orbit 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
We used XMM-Newton to observe TIC 277 for 36 ks, i.e. two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E. Ilin). We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band between 0.2-12 keV, as well as the OM using its broad band white light filter.

@@

 We extracted the events time series from EPIC using XMM SAS version 20 using the \texttt{evselect} task. We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively. We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extract a soft X-ray time series with a time binning of 200 s.

Simultaneously, we used the optical monitor (OM) aboard XMM-Newton, to monitor TIC 277 in its white light filter at $10$\,s cadence. The OM white light filter is centered on $4,060\AA$, with an equivalent width of $3,470\AA$. We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.

Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.

@@

\subsubsection{EPIC Spectra}
From the EPIC observations, we used standard spectral extraction procedure for point sources. We extracted spectra using \texttt{evselect} to select all events in the circular event and background regions,\texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra. Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin. To prduce spectra for the quiescent and flaring portions of the observations separately, we used \texttt{tabgtigen} to select the time intervals to pass to \texttt{evselect}.

@@

We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence by TESS. Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).

We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}. The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive originally masked data points following the original mask, and finally, additional 3 adjacent data point before and after the so expanded mask.

@@

\subsection{Spectral fitting: EPIC}
\label{sec:methods:epic}
We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}. We analyzed both the full data set, and the quiescent and flaring parts separately. Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model either. In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase. From the same fit, but using PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band. For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest of the upper and lower uncertainties as standard deviation.

@@

For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}. We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}. From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:

\begin{equation}
    f_{\rm OM} = \frac{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})}  d\lambda}{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})} d\lambda} 
    \label{eq:ratio}
\end{equation}

@@

Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the TESS response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of TESS, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}. Besides the $ED$, \texttt{altaipony} also yields flare start and end times, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitudes $a$.

@@

From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}). Using the combined quiescent and flaring data, we found a two-temperature VAPEC model to fit better than a one-temperature solution and no worse than a three-temperature solution. The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively. With these temperatures as ground truth, we also fitted two-temperature VAPEC models to the flaring and quiescent portions of the observations separately. The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall. However, the emission measure weighted mean temperature change dfrom \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum. This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona. In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).

Not using the full set of observations as ground truth resulted in an unconstrained hot coronal component. This can be interpreted as due to a lack of X-ray flux above 5 keV. Hot flares with temperatures of >50 MK have been measured on DS Tuc A~\citep{pillitteri2011xray}. On the Sun, the highest flare associated coronal temperatures are around 35 MK~\citep{kay2003soft}.

From the PN spectra, we measured a soft X-ray flux of \FX, and a luminosity \LX in the $0.2-2\,$keV band. 
 \begin{table}
% \movetableright=-20mm
\centering
    \script{TABLE_specfits.py}
    \caption{XSPEC fits to EPIC spectra for different subsets of observations.}
    \input{output/mcmc_specfit.tex}
        \label{tab:specfit}
\end{table}

\begin{figure}
    \script{FIGURE_data_resid.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/joint_chain_fit_data_resid.png}
        \caption{
         Best-fit joint Double APEC spectrum fit with Fe/0$=0.6$.
        }
        \label{fig:spec_joint_all}
    \end{centering}
\end{figure}

@@

\begin{table}
% \movetableright=-20mm
\centering
    \script{TABLE_tess_flares.py}
    \caption{Flares detected with TESS. $a$ is the relative amplitude, and $E_{\rm bol}$ is the bolometric flare energy assuming a 10,000 K blackbody emission from the flare.}
    \input{output/tess_flares.tex}
        \label{tab:flares}
\end{table}

@@

The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha. The frequency $R_{31.5}$ of flares per day above $\log_{10} E = 31.5\,\mathrm{erg}$ is about $\log_{10}R_{31.5}=$\ffdr. The slope is typical of other flaring stars, regardless of spectral type a rotation period~(see, e.g., Fig.~13 in \citealt{ilin2021flares}). $R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.

For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-5\,$keV band. The energy of the corresponding OM flare is \eom, which fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit~(Fig.~\ref{fig:ffd}). About one flare every $10\,$h of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277, assuming all optical flares above this energy in the OM data correspond to X-ray flares of similar or higher energy to the observed. \citet{guarcello2019simultaneous} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled energies are above $10^{33}\,$erg both in X-ray and optical. \citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result. We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely eclipsed by the star~\citep{johnstone2012soft}.

\begin{figure}
    \script{FIGURE_tess_ffd.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/277539431_tess_ffd.png}
        \caption{
         Cumulative flare frequency distribution and power law fit.
        }
        \label{fig:ffd}
    \end{centering}
\end{figure}

@@

\subsection{Flaring activity -- a young TRAPPIST-1}
\label{sec:discussion:flares}

$\log_{10}(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca. $-1.1$, consistent with other saturated late M dwarfs~\citep{medina2020flare, murray2022study}. It is similar to TRAPPIST-1's flaring rate~(XXXX), which has been investigated for its effects on the seven terrestrial planets in its orbit.

>>>>

% Title
\title{The corona of a fully convective star with a near-polar flare}

@@

% Abstract with filler text
\begin{abstract}
    In 2020, the Transiting Exoplanet Survey Satellite (TESS) observed a rapidly rotating M7 dwarf produce a flare at 81 deg latitude, the highest latitude flare located to date. This is in stark contrast to solar flares that occur much closer to the equator, typically below 30 deg. The circumstances that allow flares at high latitudes to occur are poorly understood. We studied five Sectors of TESS monitoring, and obtained 36 ks of XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to coronal temperatures and flaring activity, but is underluminous in soft X-ray emission by about an order of magnitude. Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277.

    %Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band. In combination with the TESS data, we infer that TIC 277 produces flares at or above this energy approximately once a day. 
    %Our results indicate that TIC 277 is underluminous in X-ray copX-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
Most M dwarfs hosts terrestrial planets, with a sizable fraction orbiting at instellations where liquid water can exist~\citep{dressing2015occurrence, hardegree-ullman2019kepler, ment2023occurrence}. However, to be habitable in an Earth-like manner, merely the possibility of surface water in its liquid form does not suffice. Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere affects hospitality for life, for M dwarfs in particular~\citep{airapetian2020impact}. If the energetic (particle) flux is too high, the atmosphere may blow off~\citep[e.g., ][]{lammer2003atmospheric, garraffo2017threatening, ketzer2022influence} and water may be lost from the surface~\citep{doamaral2022contribution}. If not energetic enough, life may not emerge in the first place~\citep{rimmer2018origin}.

In M dwarfs, stellar activity evolution on the main sequence unfolds much slower than for an FGK star. While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{johnstone2021active, medina2022galactic}. Moreover, the zone where liquid water can reside is much closer to these faint stars. The result is a planet that may be exposed to violent space weather conditions for a good fraction of the age of the universe.

Stellar space weather originates from the stellar corona. The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission. The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~\citep{zarka1998auroral,pineda2017panchromatic}. Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{hambaryan2004new, gizis2013kepler, paudel2020k2, schmidt2019largest}. How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{mullan2018frequencies}.%. An intermittent corona is conceivable, similarly XXXX. Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within, and consequently influences the complex process of flux emergence

\citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with TESS. They found four flares on four rapidly rotating ($P_{\rm rot}<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star. The flares were observed to rotate in and out of view. The shape of this modulation together with known inclination (combining $P_{\rm rot}$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares. Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude. Their results indicate a preference for flares at high latitudes in these stars, as a chance finding was unlikely ($\sim 0.1\%$). If flares and associated particle eruptions indeed have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought. %Are high flaring latitudes  an effect of the dynamo at high rotation speeds? Or is there something special about these stars' coronae that allowed them to produce flares at these latitudes?

@@

TIC 277539431, TIC 277 for short, showed the highest latitude flare known to date at $\sim81\,$deg~\citep[][Table \ref{tab:starparams}]{ilin2021giant}. It is an M7 dwarf with an extremely short rotation period of $4.56\,$h. While the fast rotation and late spectral type could be indicative of a brown dwarf, the detection of flares suggests the presence of a stellar corona. %The star's age is unknown. \citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would, if not as high, be compatible with the NUV/FUV emission from younger open cluster members. 
%Is its high flaring latitude an effect of the dynamo at high rotation speeds? Or is there something special about this star's corona that led it to produce are flare at these latitudes?

In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from TESS~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}). We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, investigate scenarios that are consistent with high latitude flare occurrence ~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.

@@

\begin{table}
% \footnotesize
\centering
    \caption{TIC 277539431 (WISEA J105515.71-735611.3)}
    \begin{tabular}{ll}\hline 
         Parameter & Value  \\\hline

         % $G$ mag & $14.7189 \pm 0.0011$ [4]\\
         % $G_{BP}$ mag & \\
         % $G_{BP}$ mag & \\
         % $K_s$ mag & $9.666 \pm 0.024$ [3]\\
         % $J$ mag & $10.630 \pm 0.023$ [3] \\
         Spectral type & M7 [1]\\
         Distance $d$ & $13.70\pm0.11\,$pc [2] \\
         Effective temperature $T_{\rm eff}$ & $2680^{60}_{-50}\,$ K [3]\\
         % age & \\
         Rotation period $P_{\rm rot}$ & $273.618 \pm 0.007\,$min [1]\\
         Projected rotation velocity $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
         Inclination $i$ & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
         Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\
         Bol. luminosity $L_{\rm bol}$ & \Lbol \\\hline

@@

\subsection{XMM-Newton}
\label{sec:obs:xmm}
XMM-Newton is an X-ray telescope, which launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
We used XMM-Newton to observe TIC 277 for 36 ks, i.e. two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E. Ilin). We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band from $0.2$ and up to $12\,$keV, as well as the OM using its broad band white light filter.

@@

Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve. We extracted the events time series from EPIC using XMM SAS version 20~\footnote{"Users Guide to the XMM-Newton Science Analysis System", Issue 18.0, 2023 (ESA: XMM-Newton SOC)} using the \texttt{evselect} task. We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively. We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extract a soft X-ray time series with a time binning of $200\,$s.

Simultaneously, we used the optical monitor (OM) aboard XMM-Newton, to monitor TIC 277 in its white light filter at $10$\,s cadence. The OM white light filter is centered on $4,060\,$\AA, with an equivalent width of $3,470\,$\AA. We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.

@@

\subsubsection{EPIC Spectra}
From the EPIC observations, we used the standard spectral extraction procedure for point sources. We extracted spectra using \texttt{evselect} to select all events in the circular event and background regions, \texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra. Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin. To prduce spectra for the quiescent and flaring portions of the observations separately, we used \texttt{tabgtigen} to select the time intervals to pass to \texttt{evselect}.

@@

We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence. Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).

We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}. The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking a. outliers above $>2.5\sigma$, b. the six-fold of consecutive originally masked data points following the original mask, and c. additional three adjacent data points before and after the so expanded mask.

@@

\subsection{Spectral fitting: EPIC}
\label{sec:methods:epic}
We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}. We analyzed both the full data set, and the quiescent and flaring parts separately. Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model. In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase. From the same fit, but using PN data only, we derived the flux $F_X$ and X-ray luminosity $L_X$ in the soft $0.2-2\,$keV band. For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest difference between the 50th, and 16th and 84th percentiles as standard deviation.

@@

For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}. We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}. From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $T_{\rm flare}=10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:

\begin{equation}
    f_{\rm OM} = \frac{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})}  d\lambda}{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm flare})} d\lambda} 
    \label{eq:ratio}
\end{equation}

@@

Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the TESS response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of TESS, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}. Besides the $ED$, \texttt{AltaiPony} also yields start and end time, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitude $a$, for each flare.

@@

From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}). The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively. Assuming these temperatures as ground truth, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately. The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall. However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum. From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X$\FX, and a luminosity $L_X=$\LX in the $0.2$-$2\,$keV band. %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona. In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).

We note that not using the full set of observations as ground truth resulted in an unconstrained hot coronal component. This can be interpreted as due to a lack of X-ray flux above $5\,$keV. Hot flares with temperatures of $>50\,$MK have been measured on DS~Tuc~A~\citep{pillitteri2011xray}. On the Sun, the highest flare associated coronal temperatures are around $35\,$MK~\citep{kay2003soft}.

 \begin{table}
% \movetableright=-20mm
\centering
    \script{TABLE_specfits.py}
    \caption{XSPEC fits to EPIC spectra for different subsets of observations.}
    \input{output/mcmc_specfit.tex}
        \label{tab:specfit}
\end{table}

\begin{figure}
    \script{FIGURE_data_resid.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/joint_chain_fit_data_resid.png}
        \caption{
         Best-fit joint two-temperature APEC spectral fit with Fe/O$=0.6$. The top panel shows the spectra taken with the different EPIC instruments. The bottom panel shows the residuals to the fit.
        }
        \label{fig:spec_joint_all}
    \end{centering}
\end{figure}

@@

\begin{table}
% \movetableright=-20mm
\centering
    \script{TABLE_tess_flares.py}
    \caption{Flares detected with TESS. $t_s$ is the starting time of the flare, $a$ is its relative amplitude, and $E_{\rm bol}$ is the bolometric energy assuming a 10,000 K blackbody emission from the flare.}
    \input{output/tess_flares.tex}
        \label{tab:flares}
\end{table}

@@

The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha\unskip. The frequency $R_{31.5}$ of flares per day above $\log_{10} E_{\rm flare} = 31.5\,\mathrm{erg}$ is about $\log_{10}R_{31.5}=$\ffdr\unskip. The slope is typical of other flaring stars, regardless of spectral type and rotation period~(see, e.g., Fig.~13 in \citealt{ilin2021flares}). $R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.

For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-5\,$keV band. The energy of the corresponding OM flare is \eom\unskip. We use the 36ks observing baseline of XMM-Newton to calculate a flare rate in OM. Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of TESS' flare frequency distributions to lower energies. About one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277. However, this assumes that all optical flares above this energy in the optical data correspond to detected X-ray flares in the EPIC instruments, which is not always the case~\citep{paudel2021simultaneous}. \citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled flare energies are two orders of magnitude above the flare in this work, both in X-ray and optical. \citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result. We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely eclipsed by the star~\citep{johnstone2012soft}.

\begin{figure}
    \script{FIGURE_tess_ffd.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/277539431_tess_ffd.png}
        \caption{
         Cumulative flare frequency distribution of TESS flares (black dots), with a power law fit (green line), and uncertainties (dotted lines). The rate of OM flares (gray square) is calculated from the observing baseline of XMM-Newton. 
        }
        \label{fig:ffd}
    \end{centering}
\end{figure}

@@
